### 2021-07-17 chapters/11_conclusion.tex
wip restructure

\chapter{Conclusion}
\label{ch:conclusion}
Our research started with comparing three mutation testing tools for Java and found that the state of the art was outdated.
We replicated the state of the art research with current data available on the mutation testing tools.
The state of art showed that two out of three mutation testing tools were not actively maintained anymore.
The mutation testing tool PIT scored best on the metrics measured in our replication study, this is also the tool we decided to use for our research.
\newline
We created a list of characteristics based on existing research. 
We then extended this list by identifying characteristic with logic and data available.
We used the identified characteristics to represent the mutants.
\newline
We clustered the mutants with two different clustering techniques; hierarchical clustering and by training a machine learning clustering model.
The hierarchical clustering showed that we can cluster mutants while maintaining effectiveness.
We achieved reductions of up to 75\% of the total amount of mutants without significantly reducing the accuracy.
\newline
The machine learning model showed a significant reduction in effectiveness.
We trained the machine learning model with parameters that were optimal according to existing research. 
We also used a machine learning model, specialized in calculating the number of clusters for fuzzy clustering, to calculate the number of clusters.
While the minimum clustering accuracy was higher than that of the hierarchical clustering, the maximum cluster accuracy was lower than that of the hierarchical clustering.
The cluster accuracy had a larger impact on the clusters of the machine learning model because there was a relatively small number of clusters.
Our research proves that it is possible to determine the means of domains as the centre of a cluster within a programming language without narrowing down the scope to specific pieces of code.
It can be applied to all Java projects that can be run with PIT. 
Which is a larger scope than in existing mutation clustering research.

\section{Future work}
\label{sec:future_work}
\subsubsection{Optimizing parameters of machine learning model}
We trained the machine learning model with specific parameters guided by science. 
These proved to be non optimal for our purpose.
This research can be extended by finding the optimal parameters for the \acrshort{som} clustering algorithm.
The results showed a higher minimum accuracy which can indicate that there is potential for clustering the mutants with the machine learning model.
Further research and fine-tuning is required to find out if the model can be trained to achieve more accurate scores.

@@

\subsubsection{Training a single machine learning model}
For this researched we trained a model for each source project. 
A different way is to train one machine learning model with the total set of mutants.
The model could be continuously trained each time a new source is added.
Further research is recommended in this direction as it will improve the applicability of this research.

\subsubsection{Different sets of characteristics}
We have proven that this set of characteristics can achieve accurate results in combination with hierarchical clustering.
It would be valuable to find out if it possible to reduce or expand the set of characteristics and still maintain accurate results.
Further experimentation can be done with different sets of characteristics.
\end{document}

>>>>

\chapter{Conclusion}
\label{ch:conclusion}
Our research started with comparing three mutation testing tools for Java and found that the state of the art was outdated.
We replicated the state of the art research with current data available on the mutation testing tools.
The state of art showed that two out of three mutation testing tools were not actively maintained anymore.
The mutation testing tool PIT scored best on the metrics measured in our replication study, this is also the tool we decided to use for our research.
\newline
We created a list of identified characteristics based on existing research. 
We then extended this list by identifying characteristic with logic and data available.
We used the identified characteristics to represent the mutants.
\newline
We clustered the mutants with white box approach and a black box approach.
For our white box approach we selected the hierarchical clustering algorithm.
The hierarchical clustering showed that we can cluster mutants while maintaining effectiveness.
We achieved reductions of up to 75\% of the total amount of mutants without significantly reducing the accuracy.
\newline
For the black box approach we selected a machine learning model, specifically the \acrlong{fcm} model.

The \acrshort{fcm} model showed a significant reduction in effectiveness.
We trained the\acrshort{fcm} model with parameters that were optimal according to existing research. 
We used a different machine learning model, specialized in calculating the number of clusters for fuzzy clustering, to calculate the number of clusters.
While the minimum clustering accuracy was higher than that of the hierarchical clustering, the maximum cluster accuracy was lower than that of the hierarchical clustering.
The cluster accuracy had a larger impact on the clusters of the machine learning model because there was a relatively small number of clusters.
Our research proves that it is possible to determine the means of domains as the centre of a cluster within a programming language without narrowing down the scope to specific pieces of code.
The centroids our research identified is decided by the characteristics of the mutants.
It can be applied to all Java projects that can be run with PIT. 
Which is a larger scope than in existing mutation clustering research.

\section{Future work}
\label{sec:future_work}
\subsubsection{Optimizing parameters of machine learning model}
We trained the machine learning model with specific parameters guided by relevant research. 
These proved to be non optimal for our purpose.
This research can be extended by finding the optimal parameters for the \acrshort{som} clustering algorithm.
The results showed a higher minimum accuracy which may indicate that there is potential for clustering the mutants with the machine learning model.
Further research and fine-tuning is required to find out if the model can be trained to achieve more accurate scores.

@@

\subsubsection{Training a single machine learning model}
For this researched we trained a model for each source project. 
A different way is to train a machine learning model with the combined set of mutants.
The model could be continuously trained each time a new source is added.
Further research is recommended in this direction as it will improve the applicability of this research.

\subsubsection{Different sets of characteristics}
We have proven that this set of characteristics can achieve accurate results in combination with hierarchical clustering.
It would be valuable to find out if it possible to reduce or expand the set of characteristics and still maintain accurate results.
Further experimentation may be done with different sets of characteristics.
\end{document}

### 2021-07-17 chapters/7_experiment_design.tex
wip restructure

\documentclass[../../main]{subfiles}
\begin{document}
% Because we use sections but these sections should be interpreted as chapters for rq1 and rq2 sectioning.
\clearpage
\section{Experiment design}
In this chapter we design an experiment to validate our research.
We hypothesise that we can cluster mutants with the set of characteristics we identified
while maintaining effectiveness and reducing the amount executed when executing one mutant from each cluster that is randomly selected.
Our goal is to achieve a weighted mutation score that is as close as possible to the mutation score of a full set of executed mutants.

\subsection{Project selection}
\label{ch:project_selection}
We choose three main requirements for selecting software projects; the projects should have a test suite, the test suite should not contain failing tests and the mutation testing tool should be able to execute mutants for the sample project.
We selected projects that were also used in other research within the mutation testing and testing domain\cite{Pizzoleto2019,Yu2019PossibilityScope,Wei2021SpectralTesting, Zhang2019PredictiveTesting, Chen2018SpeedingStudy, Laurent2017AssessingPIT}.
\newline
We extend our sample by selecting projects from the first six pages of the most popular Java projects on GitHub\footnote{\url{https://github.com/search?p=7&q=language\%3Ajava+stars\%3A\%3E10000&type=Repositories}}.
The unfiltered sample contains 50+ projects.
From there we filter out all the projects that contain failing tests and that are not libraries or applications.
The result consists of a sample with fifteen projects in total.

\subsection{Baseline measurement}
The baseline for our experiment is the mutation score of the set of all mutants generated by the selected mutation testing tool.
To extract this score we need to execute a full set of mutants within PIT.
For each source we have added the PIT plugin and required configuration in the \textit{pom.xml}, this code instructs PIT to generate and execute mutants with all mutators available, an example of such code can be found in Appendix \ref{ap:PIT_maven_code}.
We can then use the \acrfull{cli} and a maven command(see Listing \ref{lst:mvn-full}) to execute PIT.
\begin{lstlisting}[label=lst:mvn-full,caption=Command used to execute full set of mutants with PIT.]
mvn -U org.pitest:pitest-maven:mutationCoverage
\end{lstlisting}

\subsection{Validation}
\label{ch:exp1_validation}
The most efficient way to measure test effectiveness with mutation testing is by executing all mutants that a mutation testing tool can possibly generate.
The goal of this experiment is to reduce the amount of mutants executed while maintaining effectiveness.
To reduce the amounts executed we cluster the mutants. 
We can measure this by counting the number of clusters we generate and compare it to the number of total mutants generated by the selected tool.
\newline
To validate how effective our method of clustering is we can compare the weighted mutation score(see Chapter \ref{ch:weighted_score}) of the clustered set to the mutation score of the full set.
The closer the weighted score is to that of a full set the more effective our set of characteristics and clustering algorithm proves to be.
In other words we want achieve a mutation score that is as close as possible as to that of a full set.
We select a statistical significance level of $\leq 0.05$.
This is the conventional threshold for declaring statistical significance\cite{Kirk1996PracticalCome}.
\newline
Depending on the effectiveness of our clustering algorithm we may loose accuracy.
This can happen if a cluster contains mutants of both results. 
We can measure the accuracy inside a cluster by calculating a percentage of all mutants that have survived against the ones that have been killed in a cluster.
If the majority of the mutants in a cluster is killed then we consider that cluster to represent a killed cluster and the other way around for survived mutants.
We consider the mutants that are not in the majority of the cluster as inaccuracy.
\newline
If the weighted mutation score of our clustered set deviates more than 5\% from the score of a full set we will reject our hypothesis.

\subsection{Random mutant selection}
\label{ch:mutant_selection}
As stated in our hypothesis we randomly select a mutant from each cluster to be executed.
To validate our sample we make use of statistical hypothesis testing\cite{Emmert-Streib2019UnderstandingInference}. 
Our null hypothesis states that there is no relation between the characteristics identified and the results of a mutant.
To test this hypothesis we select an alpha value of 0.05.
To translate; if more than five percent of the values in a sample deviates more than the significance level(see Chapter \ref{ch:exp1_validation}) we cannot reject the null hypothesis.
To make our results reproducible we will select the random mutant based on a generated seed.
The seed generated for each run will be included in the results and can be found in the Github repository\cite{rbasarat-repo}.
We will repeat our experiment 30 times with 30 different seeds.
Achieving consistent results while applying random selection will contribute to the validity of the experiment.

>>>>

\documentclass[../main]{subfiles}
\begin{document}
% Because we use sections but these sections should be interpreted as chapters for rq1 and rq2 sectioning.
\clearpage
\chapter{Experiment design}
In this chapter we design an experiment to validate our research.
We hypothesise that we can cluster mutants with the set of characteristics we identified
while maintaining effectiveness and reducing the amount executed when executing one mutant from each cluster that is randomly selected.
Our goal is to achieve a weighted mutation score that is as close as possible to the mutation score of a full set of executed mutants.

\section{Project selection}
\label{ch:project_selection}
We choose three main requirements for selecting software projects; the projects should have a test suite, the test suite should not contain failing tests and the mutation testing tool should be able to execute mutants for the sample project.
We selected projects that were also used in other research within the mutation testing and testing domain\cite{Pizzoleto2019,Yu2019PossibilityScope,Wei2021SpectralTesting, Zhang2019PredictiveTesting, Chen2018SpeedingStudy, Laurent2017AssessingPIT}.
\newline
We extend our sample by selecting projects from the first six pages of the most popular Java projects on GitHub\footnote{\url{https://github.com/search?p=7&q=language\%3Ajava+stars\%3A\%3E10000&type=Repositories}}.
The unfiltered sample contains 50+ projects.
From there we filter out all the projects that contain failing tests and that are not libraries or applications.
The result consists of a sample with fifteen projects in total.

\section{Baseline measurement}
The baseline for our experiment is the mutation score of the set of all mutants generated by the selected mutation testing tool.
To extract this score we need to execute a full set of mutants within PIT.
For each source we have added the PIT plugin and required configuration in the \textit{pom.xml}, this code instructs PIT to generate and execute mutants with all mutators available, an example of such code can be found in Appendix \ref{ap:PIT_maven_code}.
We can then use the \acrfull{cli} and a maven command(see Listing \ref{lst:mvn-full}) to execute PIT.
\begin{lstlisting}[label=lst:mvn-full,caption=Command used to execute full set of mutants with PIT.]
mvn -U org.pitest:pitest-maven:mutationCoverage
\end{lstlisting}

\section{Validation}
\label{ch:exp1_validation}
The most efficient way to measure test effectiveness with mutation testing is by executing all mutants that a mutation testing tool can possibly generate.
The goal of this experiment is to reduce the amount of mutants executed while maintaining effectiveness.
To reduce the amounts executed we cluster the mutants. 
We can measure this by counting the number of clusters we generate and compare it to the number of total mutants generated by the selected tool.
\newline
To validate how effective our method of clustering is we can compare the weighted mutation score(see Chapter \ref{ch:weighted_score}) of the clustered set to the mutation score of the full set.
The closer the weighted score is to that of a full set the more effective our set of characteristics and clustering algorithm proves to be.
In other words we want achieve a mutation score that is as close as possible as to that of a full set.
We select a statistical significance level of $\leq 0.05$.
This is the conventional threshold for declaring statistical significance\cite{Kirk1996PracticalCome}.
\newline
Depending on the effectiveness of our clustering algorithm we may loose accuracy.
This can happen if a cluster contains mutants of both results. 
We can measure the accuracy inside a cluster by calculating a percentage of all mutants that have survived against the ones that have been killed in a cluster.
If the majority of the mutants in a cluster is killed then we consider that cluster to represent a killed cluster and the other way around for survived mutants.
We consider the mutants that are not in the majority of the cluster as inaccuracy.
\newline
If the weighted mutation score of our clustered set deviates more than 5\% from the score of a full set we will reject our hypothesis.

\section{Random mutant selection}
\label{ch:mutant_selection}
As stated in our hypothesis we randomly select a mutant from each cluster to be executed.
To validate our sample we make use of statistical hypothesis testing\cite{Emmert-Streib2019UnderstandingInference}. 
Our null hypothesis states that there is no relation between the characteristics identified and the results of a mutant.
To test this hypothesis we select an alpha value of 0.05.
To translate; if more than five percent of the values in a sample deviates more than the significance level(see Chapter \ref{ch:exp1_validation}) we cannot reject the null hypothesis.
To make our results reproducible we will select the random mutant based on a generated seed.
The seed generated for each run will be included in the results and can be found in the Github repository\cite{rbasarat-repo}.
We will repeat our experiment 30 times with 30 different seeds.
Achieving consistent results while applying random selection will contribute to the validity of the experiment.

### 2021-07-17 chapters/rq2/1_clustering_mutants.tex
wip restructure

\documentclass[../../main]{subfiles}
\begin{document}
\section{Experiment design}
In this chapter we design an experiment to validate our black box approach.
We hypothesise that we can train a machine learning model with the set of characteristics we identified
while maintaining effectiveness and reducing the amount executed when executing one mutant from each cluster that is randomly selected.
Our goal is to achieve a mutation score that is as close as possible to the mutation score of a full set of executed mutants.
We select a machine learning model, decide on a training strategy, train the model and validate the results.

>>>>

\documentclass[../../main]{subfiles}
\begin{document}
\section{Clustering mutants}
In this chapter devise a methodology for our black box approach.
Our goal is to achieve a mutation score that is as close as possible to the mutation score of a full set of executed mutants.
We select a machine learning model and decide on a training strategy.
